---
title: "Data Description"
warning: false
message: false
sidebar:
  contents: auto
---

Details of the data set.

## Source

The data is derived from specialist Thyroid cancer units where patients presenting with symptoms (e.g. nodules) are
assessed in greater detail and decisions are made as to whether imaging or further biopsies are required.

## Data

An ASCII CSV file (`Thy3000_DATA_LABELS_Raw.csv`) with data on 584 cases has been provided. There is no data dictionary
defining what each field is nor its type available yet.

**NB** Some of the column headers have commas (`,`) in the variable names, this can cause problems when reading CSV
files. Such strings should be double-quoted which appears to be the case but something to check carefully.

| Variable (as received)                                                                                                                                                 | Type   | Description                      | Renamed to...       |
|------------------------------------------------------------------------------------------------------------------------------------------------------------------------|--------|----------------------------------|---------------------|
| `Record ID`                                                                                                                                                            | `str`  | Unique identifier.               | `record_id`         |
| `Data Access Group`                                                                                                                                                    | `str`  | Center                           | `center`            |
| `Study ID`                                                                                                                                                             | `str`  | Study Identifiers.               | `study_id`          |
| `1.1 Date of referral`                                                                                                                                                 | `date` | Date of referral                 | `referral_date`     |
| `1.2 Which clinic was the patient recruited from?`                                                                                                                     | `str`  | Recruting clinic.                | `recruiting_clinic` |
| `If Other`                                                                                                                                                             | `str`  | Other referring clinic.          |                     |
| `1.3. The date the patient was seen in clinic`                                                                                                                         | `date` | Date patient was seen in clinic. | `clinc_date`        |
| `1.4 Referral source`                                                                                                                                                  | `str`  | Source of referral               | `referral_source`   |
| `"If Other please specify`                                                                                                                                             | `str`  | Other referring source           |                     |
| `"1.4.1 If GP was it 2-week wait referral?"`                                                                                                                           |        |                                  |                     |
| `1.5. Presentation`                                                                                                                                                    |        |                                  |                     |
| `Complete?`                                                                                                                                                            |        |                                  |                     |
| `2.1. Age of the patient when seen in clinic`                                                                                                                          |        |                                  |                     |
| `2.2. Body Mass Index of patient`                                                                                                                                      |        |                                  |                     |
| `2.3. Smoking status`                                                                                                                                                  |        |                                  |                     |
| `2.4. Previous neck irradiation`                                                                                                                                       |        |                                  |                     |
| `2.5 American Society of Anaesthesiologist (ASA) score`                                                                                                                |        |                                  |                     |
| `Complete?`                                                                                                                                                            |        |                                  |                     |
| `3.1. Presentation (choice=Neck symptoms)`                                                                                                                             |        |                                  |                     |
| `3.1. Presentation (choice=Incidental lesion on imaging)`                                                                                                              |        |                                  |                     |
| `3.1. Presentation (choice=HypErthyroidism on thyroid function test)`                                                                                                  |        |                                  |                     |
| `3.1. Presentation (choice=HypOthyroidism on thyroid function test)`                                                                                                   |        |                                  |                     |
| `3.1. Presentation (choice=Symptoms of abnormal thyroid function)`                                                                                                     |        |                                  |                     |
| `3.1. Presentation (choice=Not known)`                                                                                                                                 |        |                                  |                     |
| `3.1. Presentation (choice=Other)`                                                                                                                                     |        |                                  |                     |
| `If other please specify"`                                                                                                                                             |        |                                  |                     |
| `3.1.1 Symptomatology (choice=No symptoms)`                                                                                                                            |        |                                  |                     |
| `"3.1.1 Symptomatology (choice=Neck lump (noted by patient family or doctor))"`                                                                                        |        |                                  |                     |
| `3.1.1 Symptomatology (choice=Compressive symptoms (breathing or swallowing difficulty or voice change))`                                                              |        |                                  |                     |
| `3.1.1 Symptomatology (choice=Symptoms of thyroid dysfunction)`                                                                                                        |        |                                  |                     |
| `3.1.1 Symptomatology (choice=Other)`                                                                                                                                  |        |                                  |                     |
| `"If other please specify"`                                                                                                                                            |        |                                  |                     |
| `3.1.2 Was this nodule found incidentally on imaging?`                                                                                                                 |        |                                  |                     |
| `"If yes what imaging"`                                                                                                                                                |        |                                  |                     |
| `3.2 Clinical Assessment`                                                                                                                                              |        |                                  |                     |
| `3.2.1 Retrosternal on clinical examination`                                                                                                                           |        |                                  |                     |
| `3.2.2 Palpable lymphadenopathy`                                                                                                                                       |        |                                  |                     |
| `3.2.3 Patient perception of rapid growth of nodule`                                                                                                                   |        |                                  |                     |
| `3.4 Thyroid function tests done within 3 months of presentation to clinic`                                                                                            |        |                                  |                     |
| `3.5 Ultrasound performed`                                                                                                                                             |        |                                  |                     |
| `3.5.1 Reported maximum diameter of the largest thyroid nodule on ultrasound in millimetres`                                                                           |        |                                  |                     |
| `3.5.2 Description of thyroid nodule(s) on ultrasound`                                                                                                                 |        |                                  |                     |
| `"3.5.3 If Ultrasound performed`                                                                                                                                       |        |                                  |                     |
| `U stage reported: "`                                                                                                                                                  |        |                                  |                     |
| `"3.5.4 If Ultrasound performed`                                                                                                                                       |        |                                  |                     |
| `TIRADS reported "`                                                                                                                                                    |        |                                  |                     |
| `"3.5.5 If Ultrasound performed`                                                                                                                                       |        |                                  |                     |
| `was lymphadenopathy documented"`                                                                                                                                      |        |                                  |                     |
| `3.5.6 Elastography performed`                                                                                                                                         |        |                                  |                     |
| `3.5.7 CT neck performed`                                                                                                                                              |        |                                  |                     |
| `3.5.8 If CT neck performed reason for performing it (choice=Extrathyroid extention/local invasion)`                                                                   |        |                                  |                     |
| `3.5.8 If CT neck performed reason for performing it (choice=Retrosternal extension/plan surgical approach)`                                                           |        |                                  |                     |
| `3.5.8 If CT neck performed reason for performing it (choice=Extent of lymphadenopathy)`                                                                               |        |                                  |                     |
| `3.5.8 If CT neck performed reason for performing it (choice=Tracheal compression)`                                                                                    |        |                                  |                     |
| `3.5.8 If CT neck performed reason for performing it (choice=Not known)`                                                                                               |        |                                  |                     |
| `3.5.8 If CT neck performed reason for performing it (choice=Unrelated to thyroid pathology)`                                                                          |        |                                  |                     |
| `3.5.9 MRI neck performed`                                                                                                                                             |        |                                  |                     |
| `3.5.10 If MRI neck performed reason for performing it (choice=Extrathyroid extention/local invasion)`                                                                 |        |                                  |                     |
| `3.5.10 If MRI neck performed reason for performing it (choice=Retrosternal extension/plan surgical approach)`                                                         |        |                                  |                     |
| `3.5.10 If MRI neck performed reason for performing it (choice=Extent of lymphadenopathy)`                                                                             |        |                                  |                     |
| `3.5.10 If MRI neck performed reason for performing it (choice=Tracheal compression)`                                                                                  |        |                                  |                     |
| `3.5.10 If MRI neck performed reason for performing it (choice=Not known)`                                                                                             |        |                                  |                     |
| `3.5.10 If MRI neck performed reason for performing it (choice=Unrelated to thyroid pathology)`                                                                        |        |                                  |                     |
| `3.5.11 Iodine-123 scan performed`                                                                                                                                     |        |                                  |                     |
| `3.6. FNA of thyroid nodule performed (either at time of ultrasound or later)`                                                                                         |        |                                  |                     |
| `3.6.1 If FNA performed was Thy or Bethesda stage reported"`                                                                                                           |        |                                  |                     |
| `3.6.1.1 Thy result`                                                                                                                                                   |        |                                  |                     |
| `3.6.1.2 Bethesda result`                                                                                                                                              |        |                                  |                     |
| `3.7 Core biopsy performed`                                                                                                                                            |        |                                  |                     |
| `3.8 FNA of lymph node performed`                                                                                                                                      |        |                                  |                     |
| `3.8.1 If FNA lymph node performed result`                                                                                                                             |        |                                  |                     |
| `Complete?`                                                                                                                                                            |        |                                  |                     |
| `4.1 Initial management decision`                                                                                                                                      |        |                                  |                     |
| `4.2 Date of decision (either clinic letter or MDT date)`                                                                                                              |        |                                  |                     |
| `4.3 Date of surgery or start of other treatment (if interventional)`                                                                                                  |        |                                  |                     |
| `4.4. In case of no intervention was a routine review offered`                                                                                                         |        |                                  |                     |
| `4.4.1 If yes what was the planned interval (in weeks)`                                                                                                                |        |                                  |                     |
| `4.4.2 Was Ultrasound repeated at routine review?`                                                                                                                     |        |                                  |                     |
| `4.4.3 Was FNA repeated at routine review`                                                                                                                             |        |                                  |                     |
| `4.4.4 If routine review performed was management strategy changed at review`                                                                                          |        |                                  |                     |
| `4.4.4.1 If management strategy changed at review revised management decision`                                                                                         |        |                                  |                     |
| `"4.4.4.2 If yes reason for change in management"`                                                                                                                     |        |                                  |                     |
| `If other`                                                                                                                                                             |        |                                  |                     |
| `4.4.4.3 If no change in management decision: date last seen by 'thyroid' team`                                                                                        |        |                                  |                     |
| `"4.4.4.4 If no change in management strategy at review`                                                                                                               |        |                                  |                     |
| `confirm management plan when last seen"`                                                                                                                              |        |                                  |                     |
| `4.4.5 Patient was signposted to appropriate patient SUPPORT organisation and/or provided with written PATIENT information about thyroid nodules (including leaflets)` |        |                                  |                     |
| `Complete?`                                                                                                                                                            |        |                                  |                     |
| `5.1 Type of thyroid surgery`                                                                                                                                          |        |                                  |                     |
| `5.1.1 Lymph node dissection done at this surgery`                                                                                                                     |        |                                  |                     |
| `5.1.2 Pathology`                                                                                                                                                      |        |                                  |                     |
| `"If other cancer/other diagnosis please state"`                                                                                                                       |        |                                  |                     |
| `Complete?`                                                                                                                                                            |        |                                  |                     |

```{r}
#| label: read-data
#| eval: true
#| echo: false
#| output: false
library(dplyr)
library(ggplot2)
library(ggdark)
library(knitr)
library(tidyr)

## Two options we can either source the cleaning file...
# source("r/master.R")
## Or we can load the file that is saved at the end and use that
df <- readRDS("data/r/clean.rds")

n_observations <- df |>
  dplyr::select(study_id) |>
  dplyr::distinct() |>
  count()
n_centres <- df |>
  dplyr::select(data_access_group) |>
  dplyr::distinct() |>
  count()
```

Tabular and graphical summaries of the data provided in `Thy3000_DATA_LABELS_Raw.csv` after cleaning we can summarise
the data. In total there were `r n_observations` recorded from a total of `r n_centres`.

## Quantitative Variables

### Statistics

```{r}
#| label: quantitve-summary
#| tbl-cap: "Summary statistics for quantitative variables"
#| eval: true
#| echo: false
#| output: true


## We select the varaibles we wish to summarise and check they are numeric and within each we calculate the mean,
## standard deviation (sd), meidan, inter-quartile range (p25 and p75), minimum and maximum
df_sum <- df |>
  dplyr::select(age, bmi) |> # , nodule_maximum_diameter_mm) |>
  dplyr::summarise(dplyr::across(
    dplyr::where(is.numeric),
    list(
      Mean = ~ mean(.x, na.rm = TRUE),
      SD = ~ sd(.x, na.rm = TRUE),
      Median = ~ median(.x, na.rm = TRUE),
      "Lower IQR" = ~ quantile(.x, probs = c(0.25), na.rm = TRUE),
      "Upper IQR" = ~ quantile(.x, probs = c(0.75), na.rm = TRUE),
      Min = ~ min(.x, na.rm = TRUE),
      Max = ~ max(.x, na.rm = TRUE)
    )
  ))

## This gives us a very wide data frame so we need to reshape this using the tidyr package.
df_sum |>
  tidyr::gather(stat, val) |>
  # dplyr::mutate(stat = stringr::str_replace("nodule_maximum_diameter") ) |>
  tidyr::separate(stat, into = c("Variable", "Statistic"), sep = "_") |>
  tidyr::spread(Statistic, val) |>
  dplyr::select(Variable, Mean, SD, "Lower IQR", Median, "Upper IQR", Min, Max) |>
  knitr::kable(digits = 3)
```

### Age

```{r}
#| label: age-hist
#| fig-cap: "Distribution of Ages at referral"
#| eval: true
#| echo: false
#| output: true

df |> ggplot(aes(age)) +
  geom_histogram() +
  labs(x = "Age", y = "N") +
  dark_theme_minimal()
```

### Body Mass Index (BMI)

```{r}
#| label: bmi-hist
#| fig-cap: "Distribution of Body Mass Index (BMI)"
#| eval: true
#| echo: false
#| output: true

df |> ggplot(aes(bmi)) +
  geom_histogram() +
  labs(x = "BMI", y = "N") +
  dark_theme_minimal()
```

### Nodule Maximum Diameter

```{r}
#| label: nodule-maximum-diameter-hist
#| fig-cap: "Distribution of Nodule Maximum Diameter (mm)"
#| eval: true
#| echo: false
#| output: true

df |> ggplot(aes(nodule_maxmimum_diameter_mm)) +
  geom_histogram() +
  labs(x = "Nodule Maximum Diameter (mm)", y = "N") +
  dark_theme_minimal()
```

## Categorical Variables

### ASA Score

```{r}
#| label: asa-bar
#| fig-cap: "Reported ASA score"
#| eval: true
#| echo: false
#| output: true

df |> ggplot(aes(asa_score, fill = asa_score)) +
  geom_bar() +
  labs(x = "ASA Score", y = "N") +
  scale_fill_discrete(guide = "none") +
  dark_theme_minimal()
```

```{r}
#| label: asa-table
#| tbl-cap: "Reported ASA score"
#| eval: true
#| echo: false
#| output: true

df |>
  dplyr::count(asa_score) |>
  dplyr::mutate("%" = prop.table(n) * 100) |>
  knitr::kable(digits = 3)
```

### Data Access Group

```{r}
#| label: data-access-group-bar
#| fig-cap: "Data Access Group"
#| eval: true
#| echo: false
#| output: true

df |> ggplot(aes(data_access_group, fill = data_access_group)) +
  geom_bar() +
  labs(x = "Data Access Group", y = "N") +
  scale_fill_discrete(guide = "none") +
  dark_theme_minimal() +
  ## We rotate the text on the x-axis by 90 degrees so it is legible
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
#| label: data-access-group-table
#| tbl-cap: "Data Access Group"
#| eval: true
#| echo: false
#| output: true

df |>
  dplyr::count(data_access_group) |>
  dplyr::mutate("%" = prop.table(n) * 100) |>
  knitr::kable(digits = 3)
```

### Referral Source

#| label: referral-source-bar
#| fig-cap: "Referral Source"
#| eval: true
#| echo: false
#| output: true
df |> ggplot(aes(referral_source, fill = referral_source)) +
  geom_bar() +
  labs(x = "referral_source", y = "N") +
  scale_fill_discrete(guide = "none") +
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))
```

```{r}
#| label: referral-source-table
#| tbl-cap: "Referral Source"
#| eval: true
#| echo: false
#| output: true

df |>
  dplyr::count(referral_source) |>
  dplyr::mutate("%" = prop.table(n) * 100) |>
  knitr::kable(digits = 3)
```


### Recruiting Clinic

```{r}
#| label: recruiting-clinic-bar
#| fig-cap: "Recruiting Clinic"
#| eval: true
#| echo: false
#| output: true

df |> ggplot(aes(clinic_recruiting, fill = clinic_recruiting)) +
  geom_bar() +
  labs(x = "Recruiting Clinic", y = "N") +
  scale_fill_discrete(guide = "none") +
  dark_theme_minimal() +
  ## Here we wrap the text of the long labels
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))
```

```{r}
#| label: recruiting-clinic-table
#| tbl-cap: "Recruiting Clinic"
#| eval: true
#| echo: false
#| output: true

df |>
  dplyr::count(clinic_recruiting) |>
  dplyr::mutate("%" = prop.table(n) * 100) |>
  knitr::kable(digits = 3)
```

### Thyroid Nodule FNA

```{r}
#| label: nodule-fna-thy-bar
#| fig-cap: "Thyroid Nodule FNA"
#| eval: true
#| echo: false
#| output: true

df |> ggplot(aes(module_fna_thy, fill = module_fna_thy)) +
  geom_bar() +
  labs(x = "Thyroid Nodule FNA", y = "N") +
  scale_fill_discrete(guide = "none") +
  dark_theme_minimal() +
  ## Here we wrap the text of the long labels
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))
```

```{r}
#| label: nodule-fna-thy-table
#| tbl-cap: "Thyroid Nodule FNA"
#| eval: true
#| echo: false
#| output: true

df |>
  dplyr::count(module_fna_thy) |>
  dplyr::mutate("%" = prop.table(n) * 100) |>
  knitr::kable(digits = 3)
```


### Final Pathology

```{r}
#| label: final-pathology-bar
#| fig-cap: "Final Pathology"
#| eval: true
#| echo: false
#| output: true

df |> ggplot(aes(final_pathology, fill = final_pathology)) +
  geom_bar() +
  labs(x = "Final Pathology", y = "N") +
  scale_fill_discrete(guide = "none") +
  dark_theme_minimal() +
  ## Here we wrap the text of the long labels
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))
```

```{r}
#| label: final-pathology-table
#| tbl-cap: "Final Pathology"
#| eval: true
#| echo: false
#| output: true

df |>
  dplyr::count(final_pathology) |>
  dplyr::mutate("%" = prop.table(n) * 100) |>
  knitr::kable(digits = 3)
```


### Final Pathology x Thyroid Nodule FNA

It is useful to check that the rules we have derived for classifying Final Pathology, which are based on
`thyroid_surgery_lymph_node_pathology` when available and `nodule_fna_thy` when not are correct. We can do this by
plotting and tabulating the data.
```{r}
#| label: nodule-fna-thy-bar
#| fig-cap: "Thyroid Nodule FNA"
#| eval: true
#| echo: false
#| output: true

df |> ggplot(aes(final_pathology, fill = final_pathology)) +
  geom_bar() +
  labs(x = "Thyroid Nodule FNA", y = "N") +
  scale_fill_discrete(guide = "none") +
  dark_theme_minimal() +
  ## Here we wrap the text of the long labels
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15))
```

```{r}
#| label: surgery-pathology-nodule-fna-thy-table
#| tbl-cap: "Surgical Lymp Node Pathology and Thyroid Nodule FNA. Percentages are by Pathology"
#| eval: true
#| echo: false
#| output: true

df |>
  dplyr::group_by(thyroid_surgery_lymph_node_pathology) |>
  dplyr::count(nodule_fna_thy) |>
  dplyr::mutate(
    percent = prop.table(n) * 100,
    "N (%)" = paste0(n, " (", format(round(percent, 3), nsmall = 3), ")")
  ) |>
  ## NB - No need to include thyroid_surgery_lymph_node_pathology as we have group_by that variable
  dplyr::select(nodule_fna_thy, "N (%)") |>
  tidyr::spread(nodule_fna_thy, "N (%)") |>
  setNames(c(
    "Thyroid Surgery Lymph Node Pathology", "Thy1 (%)", "Thy2 (%)", "Thy3 (%)", "Thy4 (%)", "Thy5 (%)",
    "Missing (%)"
  )) |>
  knitr::kable(digits = 3)
```

We may be interested in the percentages across all observations though.
```{r}
#| label: surgery-pathology-nodule-fna-thy-table
#| tbl-cap: "Surgical Lymp Node Pathology and Thyroid Nodule FNA. Percentages are across all observations."
#| eval: true
#| echo: false
#| output: true

df |>
  dplyr::group_by(thyroid_surgery_lymph_node_pathology) |>
  dplyr::count(nodule_fna_thy) |>
  ## We ungroup the data to get percentages across all pathologies
  dplyr::ungroup() |>
  dplyr::mutate(
    percent = prop.table(n) * 100,
    "N (%)" = paste0(n, " (", format(round(percent, 3), nsmall = 3), ")")
  ) |>
  ## NB - We MUST now include thyroid_surgery_lymph_node_pathology as the data is not grouped
  dplyr::select(thyroid_surgery_lymph_node_pathology, nodule_fna_thy, "N (%)") |>
  tidyr::spread(nodule_fna_thy, "N (%)") |>
  setNames(c(
    "Thyroid Surgery Lymph Node Pathology", "Thy1 (%)", "Thy2 (%)", "Thy3 (%)", "Thy4 (%)", "Thy5 (%)",
    "Missing (%)"
  )) |>
  knitr::kable(digits = 3)
```
